Are environmental and geographic effective surrogates for genetic variation in conservation planning?

### 2016-01-20 article/article.Rmd
progress on manuscript

## Introduction

## Methods
### Study area
To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [Figure S1; @r451; see @r452 for methodological details]. This dataset was well suited for this study, because it provides spatially explicit genomic data for a multitude of species across a broad geographic area and over large environmental gradients. As part of this project, data for 27 alpine plant species were collected across the European Alps according to a grid (22.3 km $\times$ 25 km). Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453]. Matrices denoting the presence/absence of polymorphisms at loci were constructed independently for each species. Thus for each second grid cell, for each of the species in the grid cell, the dataset contains information describing the genomic variation of individuals inside the grid cell. We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation.

### Surrogate data
We explored the effectiveness of climatic and geographic surrogates. To describe variation in the location of each grid cell, we projected the grid into an equi-distant coordinate system (ESRI:102031) and used the centroid of each cell. To describe the climatic variation among each cell, we obtained twenty-one bioclimatic variables [approx. 1km resolution; @r33]. These layers and the grid cells were projected into an equal-area coordinate system (ESRI:102014). The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1). The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)`% of the variation and were used to generate two new layers. We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).

@@

First, we grouped individuals into genetic lineages for each species (Figure S2). The purpose of this step was to avoid confounding variation between populations attributed to genetic history instead of selection pressures. We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in \texttt{adegenet} R package; @r468]. Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca`% variation. We subject these components to multiple $k$-means clustering models with varying numbers of assumed populations (1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.inter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts). We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups. We selected the model before the first group switch, and used the population memberships in that model for subsequent analysis.

@@

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if present) and neutral genetic variation for each species (Figures S4--S29). For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457]. These distances were subject to multiple non-metric multi-dimensional scaling analyses using the minimum number of dimensions required to acheive an adequate stress value  [`r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; implemented in the \texttt{vegan} R package; @r458]. Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize the genomic variation among the grid cells.

### Prioritizations
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve a proportion of the species' range as well as a proportion of the species' variation [using an attribute space; _sensu_ @r465]. We treated the genomic and surrogate variables as distinct attribute spaces. We set the demand points for each species and its attribute space using the values associated with the planning units.

To fulfil the aims of the study, we generated three sets of prioritisations. First, we generated a set of single species prioritizations. For each species, we generated a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target, a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)`% space-based targets for the surrogate variation, and a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100) `% space-based targets for the genomic variation. Next, we generated three multi-species prioritizations using the same sets of targets used in the single species prioritizations. Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)`% and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)`% in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100`%. All of these prioritisations had no amount-based targets. For each prioritization we calculated the proportion of the adaptive (if present) and neutral genetic variation that it secured. We solved the reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)`% of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

## Statistical analysis
We tested for differences in the effectiveness of prioritisations at securing genomic variation using general linear models (GLMs). These models were fit to the proportion of genomic variation explained using a logit link. They contained a predictor variable indicating if the prioritisation was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genomic variation was adaptive or neutral, and the interaction between these variables. These models were subject to stepwise term deletion routes using likelihood ratio tests (LRT) to identify the minimal adequate model (terms were retained when $P < 0.05$).

>>>>

## Introduction
Biodiversity is under threat. Many species have gone extinct over the last few decades [@r270; @r272]. Currently there are thousands of species listed as endangered or near threatened in the International Union for Conservation of Nature (IUCN) red list (\url{www.iucnredlist.org}). Futhermore, climate change is expected cause even greater loss to global biodiversity [@r236; @r243]. In response to this crisis, the overarching aim of conservation is to maximise the long-term persistance of biodiversity [@r255; @r67].

The act of setting aside areas for preservation is the hallmark of modern conservation [@r440]. Protected areas buffer species from anthropogenic impacts and provide areas for biodiversity processes to persist (eg. nutrient cycling and water purification). Additionally, protected areas set the stage for direct management actions (eg. habitat rehabitilation, translocation, and invasive species management). Since the resources avaliable for conservation are limited, the plans for protected areas (prioritizations) are often generated using spatially explicit biological data (eg. species distribution data) and decision support tools [eg. $\texttt{MARXAN}$; @r22; $\texttt{Zonation}$; @r429] that identify cost-effective solutions.

Protected areas must preserve intra-specific genetic variation to ensure their persistance [@r67; @r3; @r61]. This variation can be classified as either adaptive or neutral [reviewed in @r????]. Adaptive genetic variation is associated with loci currently under selection, whereas neutral genetic variation is associated with loci that are not. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3]. In constrast, neutral variation reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can secure different genetic lineages and avoid the adverse effects of low genetic diversity [eg. inbreeding depression; @r67]. However, genetic data requires considerable resources to obtain and analyse, and as a consequence, conservation planners do not typically use genetic data to inform prioritizations [@r61]. Thus prioritisations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver desired outcomes.

Recently, conservation scientists have begun investigating the use of surrogates for genetic data to generate more effective prioritisations without the overheads associated with genetic data. Since adaptive genetic variation is driven by selection pressures, environmental variables have been explored as potential surrogates for genetic variation [eg. @r247; @r2]. On the other hand, neutral variation arises from a reduction in gene flow, and so variables that predict reduction in gene flow have been explored. For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have lower gene flow, and in turn have greater neutral variation among them. The use of distance-based surrogates has been found to improve prioritisations in an island network [@r200]. However the efftiveness of this surrogate remains untested in systems where connectivity is complicated by additional factors (eg. elevation in terrestrial systems). But before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates needs to be verified.

Our overarching aim is to investigate the effectiveness of environmental and geographic surrogates for conservation planning. First, we aim to test the effectiveness of these surrogates under the simplest of conditions by using a set of single-species prioritisations. Second, we aim to test the effectiveness of these surrogates in a more realistic multi-species context, and see how they compare to the single-species prioritisations. Third, we aim to understand the relationship between the proportion of surrogate variation secured in a multi-species prioritisation and the prorpotion of genetic variation secured in a prioritization. If these surrogates markedly improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

## Methods
### Study area
To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r451; see @r452 for methodological details]. This dataset was well suited for this study, because it provides spatially explicit genomic data for a multitude of species across a broad geographic area and over large environmental gradients. As part of this project, data for 27 alpine plant species were collected across the European Alps according to a grid (22.3 km $\times$ 25 km; Figure S1). Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453]. Matrices denoting the presence/absence of polymorphisms at loci were constructed independently for each species. Thus for each second grid cell, for each of the species in the grid cell, the dataset contains information describing the genomic variation of individuals inside the grid cell. We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation.

### Surrogate data
We explored the effectiveness of climatic and geographic surrogates. To describe variation in the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$) and used the centroid of each cell. To describe the climatic variation among each cell, we obtained twenty-one bioclimatic variables [approx. 1 km resolution; @r33]. These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1). The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers. We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).

@@

First, we grouped individuals into genetic lineages for each species (Figure S2). The purpose of this step was to avoid confounding variation between populations attributed to genetic history instead of selection pressures. We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in $\texttt{adegenet}$ R package; @r468]. Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca` % variation. We subject these components to multiple $k$-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts). We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recomended in the \texttt{adegenet} user manual). The populations members associated with these models were used for subsequent analysis.

@@

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if present) and neutral genetic variation for each species (Figures S4--S29). For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457]. These distances were subject to multiple non-metric multi-dimensional scaling analyses using the minimum number of dimensions required to acheive an adequate stress value  [$\leq$ `r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; implemented in the $\texttt{vegan}$ R package; @r458]. Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize the genomic variation among the grid cells.

### Prioritizations
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve a proportion of the species' range as well as a proportion of the species' variation using an attribute space [_sensu_ @r465]. We treated the genomic and surrogate variables as distinct attribute spaces. We set the demand points for each species and its attribute space using the values associated with the planning units.

To fulfil the aims of the study, we generated three sets of prioritisations. First, we generated a set of single species prioritizations. For each species, we generated `r rapr.params.LST[[MODE]]$single.species.amount.replicates` prioritizations using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target, a single prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % space-based targets for the surrogate variation, and single a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100) ` % space-based targets for the genomic variation. We generated multiple solutions for the single-species prioritizations containing only amount-based targets because they were effectively a random selection of the planning units that the species inhabits, and so we obtained multiple solutions to account for stochasticity. Next, we generated three multi-species prioritizations using the same sets of targets used in the single species prioritizations. Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100` %. All of these prioritisations had no amount-based targets. For each prioritization we calculated the proportion of the adaptive (if present) and neutral genetic variation that it secured. We solved the reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

## Statistical analysis
We tested for differences in the effectiveness of prioritizations at securing genetic variation using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469]. We fit two full models to the proportion of genomic variation secured in the single- and multi-species prioritizations using a logit link. Both models contained a predictor variable indicating if the prioritisation was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genomic variation was adaptive or neutral, and the interaction between these variables. Additionally, both models contained an intercept-only random effect to ensure that comparisons were made within species. Each full model was subject to a stepwise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$). Finally, we conducted a post-hoc analysis using Tukey constrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

### 2016-01-20 article/figures.Rmd
progress on manuscript

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    toc: false
    keep_tex: no
    fig_caption: yes
    includes:
      in_header: preamble-latex.tex
---

@@

```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritisations. Single species prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Data shows the performance of prioritisations generated using these three sets of targets. Bars denote means and standard errors.'}
# prepare data for plotting
single.spp.PDF <- expand.grid(
	Prioritisation=unique(single.spp.SDF$Prioritisation),
	Metric=unique(single.spp.SDF$Metric))
single.spp.PDF <- cbind(single.spp.PDF,
	as.data.frame(predict(single.spp.GLM, single.spp.PDF,
		type='response', se.fit=TRUE))) %>%
	mutate(lower=fit-se.fit, upper=fit+se.fit,
		letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
		letter_pos=upper+0.05)
# make plot
ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
	data=single.spp.PDF) +
	geom_bar(position=position_dodge(0.9),
		stat='identity') +
	geom_errorbar(
		aes(ymin=lower,ymax=upper),
		position=position_dodge(0.9), width=0.6) +
	geom_text(aes(x=Metric, y=letter_pos,
		label=letters), position=position_dodge(0.9)) +
	scale_fill_manual(name='Prioritisation',
		values=c('grey80','grey50','grey20')) +
	ylab('Proportion genetic\nvariation secured (%)') +
	xlab('') +
	theme_classic()
```

```{r, echo=FALSE, fig.width=8, fig.height=3.5, fig.cap='Multi-species prioritisations. Panel (a) shows the prioritisation generated for using just amount-based targets. Panel (b) shows the prioritisation generated using amount-based and surrogate based targets. Panel (c) shows the prioritisation generated using amount-based and genetic-based targets'}
# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,] %>% spFortify
# prepare data for plotting
multi.spp.grid.FPLY <- grid.PLY
for (i in seq_along(multi.spp.prioritisations))
	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
# make maps
do.call(
	grid.arrange,
	append(
		llply(
			seq_along(multi.spp.prioritisations),
			function(i) {
				ggplot() +
					geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
						fill='grey20', color='grey80') +
					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
						group='group', fill=paste0('v',i)),
						alpha=0.8, color='grey10') +
					guides(fill=guide_legend(title=' ')) +
					theme_classic() +
					theme(axis.ticks=element_blank(), axis.text=element_blank(),
						plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
						legend.position='none') +
					coord_cartesian(
						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
					) +
					xlab('') +
					ylab('') +
					annotate('text',
						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
						label=letters[i], hjust=1, vjust=1, color='white', size=8)
			}
		),
		list(nrow=1)
	)
)
```

```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritisations. Three prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Data shows the performance of these prioritisations based on how much genetic variation they explain. Bars denote means and standard errors.'}
# prepare data for plotting
multi.spp.PDF <- expand.grid(
	Prioritisation=unique(multi.spp.SDF$Prioritisation),
	Metric=unique(multi.spp.SDF$Metric))
multi.spp.PDF <- cbind(multi.spp.PDF,
	as.data.frame(predict(multi.spp.GLM, multi.spp.PDF,
		type='response', se.fit=TRUE))) %>%
	mutate(lower=fit-se.fit, upper=fit+se.fit,
		letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
		letter_pos=upper+0.05)
# make plot
ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
	data=multi.spp.PDF) +
	geom_bar(position=position_dodge(0.9),
		stat='identity') +
	geom_errorbar(
		aes(ymin=lower,ymax=upper),
		position=position_dodge(0.9), width=0.6) +
	geom_text(aes(x=Metric, y=letter_pos,
		label=letters), position=position_dodge(0.9)) +
	scale_fill_manual(name='Prioritisation',
		values=c('grey80','grey50','grey20')) +
	ylab('Proportion genetic\nvariation secured (%)') +
	xlab('') +
	theme_classic()
```

```{r, echo=FALSE, fig.width=8, fig.height=3.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.'}
# make plots
p1 <- ggplot(data=env.pareto.DF) +
	geom_line(aes(x=Surrogate.target,y=adaptive.held,group=Species),
		alpha=0.5) +
	xlab('Environmental variation secured (%)') +
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic()
p2 <- ggplot(data=geo.pareto.DF) +
	geom_line(aes(x=Surrogate.target,y=neutral.held,group=Species),
		alpha=0.5) +
	xlab('Geographic variation secured (%)') +
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic()
grid.arrange(p1, p2, nrow=1)
```

>>>>

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---

@@

```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritizations. Single species prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Violin plots show the performance of prioritizations generated using these three sets of targets. Points denote the values associated with each prioritization. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
single.spp.PDF <- expand.grid(
	Prioritisation=unique(single.spp.SDF$Prioritisation),
	Metric=unique(single.spp.SDF$Metric)
) %>% mutate(
	letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
	letter.height=max(single.spp.SDF$value)+(diff(range(single.spp.SDF$value))*0.1)
)
# make plot
ggplot(aes(x=Metric,y=value,fill=Prioritisation),
	data=single.spp.SDF) +
	geom_violin(position=position_dodge(0.9)) +
	geom_point(alpha=0.2,position=position_jitterdodge(dodge.width=0.9, jitter.width=0.25)) +
	geom_text(aes(x=Metric, y=letter.height, label=letters),
		data=single.spp.PDF, position=position_dodge(0.9)) +
	scale_fill_manual(name='Prioritisation',
		values=c('grey80','grey50','grey20')) +
	ylab('Genetic variation secured (%)') +
	xlab('') +
	theme_classic()
```

```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='Multi-species prioritizations. Panel (a) shows the prioritization generated for using just amount-based targets. Panel (b) shows the prioritization generated using amount-based and surrogate based targets. Panel (c) shows the prioritization generated using amount-based and genetic-based targets.'}
# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,] %>% spFortify
# prepare data for plotting
multi.spp.grid.FPLY <- grid.PLY
for (i in seq_along(multi.spp.prioritisations))
	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
# make maps
do.call(
	grid.arrange,
	append(
		llply(
			seq_along(multi.spp.prioritisations),
			function(i) {
				ggplot() +
					geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
						fill='grey20', color='grey80') +
					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
						group='group', fill=paste0('v',i)),
						alpha=0.8, color='grey10') +
					guides(fill=guide_legend(title=' ')) +
					theme_classic() +
					theme(axis.ticks=element_blank(), axis.text=element_blank(),
						plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
						legend.position='none') +
					coord_cartesian(
						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
					) +
					xlab('') +
					ylab('') +
					annotate('text',
						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
						label=letters[i], hjust=1, vjust=1, color='white', size=8)
			}
		),
		list(nrow=1)
	)
)
```

```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritizations. Three prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Data shows the performance of these prioritizations based on how much genetic variation they explain. See Figure 1 caption for conventions.'}
# prepare data for plotting
multi.spp.PDF <- expand.grid(
	Prioritisation=unique(multi.spp.SDF$Prioritisation),
	Metric=unique(multi.spp.SDF$Metric)
) %>% mutate(
	letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
	letter.height=max(multi.spp.SDF$value)+(diff(range(multi.spp.SDF$value))*0.1)
)
# make plot
ggplot(aes(x=Metric,y=value,fill=Prioritisation),
	data=multi.spp.SDF) +
	geom_violin(position=position_dodge(0.9)) +
	geom_point(alpha=0.2,position=position_jitterdodge(dodge.width=0.9, jitter.width=0.25)) +
	geom_text(aes(x=Metric, y=letter.height,label=letters),
		data=multi.spp.PDF, position=position_dodge(0.9)) +
	scale_fill_manual(name='Prioritisation',
		values=c('grey80','grey50','grey20')) +
	ylab('Genetic variation secured (%)') +
	xlab('') +
	theme_classic()
```

```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Each panel shows the relationship for a different type of genetic variation. Each line denotes a different species. Lines are transparent and so darker regions show areas where species have similar relationships.'}
# make plots
p1 <- ggplot(data=env.pareto.DF) +
	geom_line(aes(x=Surrogate.target,y=adaptive.held,group=Species),
		alpha=0.3) +
	xlab('Environmental variation secured (%)') +
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic()
p2 <- ggplot(data=geo.pareto.DF) +
	geom_line(aes(x=Surrogate.target,y=neutral.held,group=Species),
		alpha=0.3) +
	xlab('Geographic variation secured (%)') +
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic()
grid.arrange(p1, p2, nrow=1)
```

### 2016-01-20 article/supporting_information.Rmd
progress on manuscript

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---
## Supporting Information

@@

```{r, fig.width=6.5, fig.height=3.25, fig.cap='Climatic variation. Each panel depicts variation based on a different principle component (PC). Squares represent planning units. The color of each planning unit denotes the average PC value of pixels inside it. Planning units with more similar colors have more similar climate regimes.', echo=FALSE}
do.call(
	grid.arrange,
		append(
		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
			ggplot() +
				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
					fill='grey20', color='grey80') +
				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
					group='group', fill=x),
					alpha=0.8, color='grey10') +
				guides(fill=guide_legend(title=' ')) +
				theme_classic() +
				theme(axis.ticks=element_blank(), axis.text=element_blank(),
					plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
				coord_cartesian(
					xlim=buffered.range(grid.FPLY$long, 0.05),
					ylim=buffered.range(grid.FPLY$lat, 0.05)
				) +
				xlab('') +
				ylab('') +
				ggtitle(paste0('PC ', substr(x, nchar(x), nchar(x))))
		}),
		list(ncol=2)
	)
)
```

@@

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}. See Figure S6 caption for conventions.')}
if (4 <= general.params.LST[[MODE]]$n.spp) if (4 <= ) plot.spp.mds(4)
```

>>>>

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---
## Supporting Information
All data and code is stored in an online repository (\url{www.github.com/paleo13/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.

@@

```{r, fig.width=8, fig.height=3.5, fig.cap='Climatic variation. Each panel depicts variation based on a different principle component (PC). Squares represent planning units. The color of each planning unit denotes the average PC value of pixels inside it. Planning units with more similar colors have more similar climate regimes.', echo=FALSE}
do.call(
	grid.arrange,
		append(
		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
			ggplot() +
				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
					fill='grey20', color='grey80') +
				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
					group='group', fill=x),
					alpha=0.8, color='grey10') +
				guides(fill=guide_legend(title=' ')) +
				theme_classic() +
				theme(axis.ticks=element_blank(), axis.text=element_blank(),
					plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
				coord_cartesian(
					xlim=buffered.range(grid.FPLY$long, 0.05),
					ylim=buffered.range(grid.FPLY$lat, 0.05)
				) +
				xlab('') +
				ylab('') +
				ggtitle(paste0('PC ', substr(x, nchar(x), nchar(x))))
		}),
		list(ncol=2)
	)
)
```

@@

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}. See Figure S6 caption for conventions.')}
if (4 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(4)
```
